Figure 5 - IP-MS

### 2024-07-09 ms_figure_5_ipms.Rmd
Fix heading Figure 5
Lift out IP-MS heatmap functions into scripts, move HOT IP-MS heatmap to Supp Fig 2

@@

# Helper functions

We define a number of additional helper functions for data extraction and 
summarization, which will be used below.

```{r heatmap-helper}
filterForHeatmap <- function(tstatmat, adjpmat, logfcmat, adjpThreshold, 
                             logfcThreshold, maxNbrTargets) {
    ## Filter - require significant in at least one comparison
    tstatfilt <- tstatmat
    tstatfilt[adjpmat >= adjpThreshold] <- NA
    tstatfilt[logfcmat <= logfcThreshold] <- NA
    # tstatfilt[tstatfilt <= 0] <- NA

    ## Some experiments pull down a large number of proteins
    # print(table(colSums(!is.na(tstatfilt))))
    ## -> remove proteins that are only being pulled down in these experiments
    keep <- which(colSums(!is.na(tstatfilt)) <= maxNbrTargets)
    tstatfilt <- tstatfilt[rowSums(!is.na(tstatfilt[, keep])) > 0, ]
    tstatfilt[is.na(tstatfilt)] <- 0
    tstatfilt <- tstatfilt[rowSums(tstatfilt) > 0, ]

    ## Subset the t-statistic matrix to only the retained proteins
    stopifnot(colnames(tstatfilt) == colnames(tstatmat))
    tstatmat <- tstatmat[rownames(tstatfilt), colnames(tstatfilt)]
    colnames(tstatmat) <- .getSimplifiedComparison(colnames(tstatmat))

    list(tstats = tstatmat, tstatsfilt = tstatfilt)
}

makeComplexAnnotation <- function(tstatmat, complexes, idmap,
                                  show_legend = TRUE) {
    rowannot <- data.frame(Complex = rep(NA, nrow(tstatmat)),
                           row.names = rownames(tstatmat))
    rowcols <- list(Complex = c())
    for (cmplx in names(complexes)) {
        rowannot$Complex[.getPomBaseIdFromProtein(rownames(rowannot), 
                                                  idmap = idmap) %in% 
                             complexes[[cmplx]]$pombase_ids] <- cmplx
        rowcols$Complex[cmplx] <- complexes[[cmplx]]$color
    }
    rowAnnotation(df = rowannot, 
                  col = rowcols, 
                  annotation_name_gp = gpar(fontsize = fl),
                  simple_anno_size_adjust = TRUE,
                  na_col = bg_color,
                  show_annotation_name = FALSE,
                  show_legend = show_legend,
                  annotation_legend_param = list(title_gp = gpar(fontsize = fl)))
}

makeComplexAndDBDAnnotation <- function(tstatmat, complexes, dbd, idmap) {
    rowannot <- data.frame(Complex = rep(NA, nrow(tstatmat)),
                           row.names = rownames(tstatmat))
    rowcols <- list(Complex = c())
    for (cmplx in names(complexes)) {
        rowannot$Complex[.getPomBaseIdFromProtein(rownames(rowannot), 
                                                  idmap = idmap) %in% 
                             complexes[[cmplx]]$pombase_ids] <- cmplx
        rowcols$Complex[cmplx] <- complexes[[cmplx]]$color
    }
    rowannot$`DBD fam.` <- dbd$DBD_class[match(
        .getPomBaseIdFromProtein(rownames(tstatmat), idmap = idmap), 
        dbd$PomBaseID)]
    setfam <- c("KilA-N", "Zn(II)2Cys6", "bZIP")
    unsetfam <- setdiff(unique(dbd$DBD_class), setfam)
    rowannot$`DBD fam.`[rowannot$`DBD fam.` %in% unsetfam] <- "Other (combined)"
    rowcols <- c(rowcols, list(`DBD fam.` = c(`KilA-N` = main_colors[5], 
                                              `Zn(II)2Cys6` = main_colors[4],
                                              bZIP = main_colors[3], 
                                              `Other (combined)` = bg_color)))
    rowAnnotation(df = rowannot, 
                  col = rowcols, 
                  annotation_name_gp = gpar(fontsize = fl),
                  simple_anno_size_adjust = TRUE,
                  na_col = bg_color,
                  show_annotation_name = TRUE,
                  show_legend = FALSE,
                  annotation_legend_param = list(title_gp = gpar(fontsize = fl)))
}

makeColLabels <- function(tstatmat, colLabel, fontsize) {
    columnAnnotation(
        TFnames = anno_mark(which = "column", side = "bottom",
                            at = match(colLabel, colnames(tstatmat)), 
                            labels = .getProteinNameFromSimplifiedComparison(colLabel),
                            labels_gp = gpar(fontsize = fontsize)))
}

makeRowLabels <- function(tstatmat, rowLabel, fontsize) {
    rowAnnotation(
        TFnames = anno_mark(which = "row", side = "left",
                            at = match(rowLabel, rownames(tstatmat)), 
                            labels = rowLabel,
                            labels_gp = gpar(fontsize = fontsize)))
}

makeHeatmapCol <- function(tstatmat, stringency = "high") {
    if (stringency == "high") {
        circlize::colorRamp2(
            breaks = seq(3.0, 13.0, length.out = 64),
            colors = colorRampPalette(binary_heatmap_colors[c("FALSE", "TRUE")])(64))
    } else {
        circlize::colorRamp2(
            breaks = seq(1.5, 13.0, length.out = 64),
            colors = colorRampPalette(binary_heatmap_colors[c("FALSE", "TRUE")])(64))
    }
}

reorderColsByRows <- function(tstat) {
    ## Row names should be protein names, column names simplified comparison names
    ord <- match(rownames(tstat), .getProteinNameFromSimplifiedComparison(colnames(tstat)))
    ord <- ord[!is.na(ord)]
    ord <- c(ord, setdiff(seq_len(ncol(tstat)), ord))
    tstat[, ord]
}

makeHeatmapData <- function(sce, adjpthr, log2fcthr, conc) {
    tc <- .getTestCols(sce, adjp_cutoff = adjpthr, logfc_cutoff = log2fcthr)

    tstats <- filterForHeatmap(tstatmat = tc$tstat, adjpmat = tc$adjp, 
                               logfcmat = tc$logfc, logfcThreshold = log2fcthr,
                               adjpThreshold = adjpthr, maxNbrTargets = Inf)
    tstat <- tstats$tstats

    ## tstats$tstatsfile: protein x experiment
    ## Number of TFs that are pulled down by each bait
    all_tfs <- setdiff(.getProteinNameFromComparison(colnames(tstats$tstatsfilt)), 
                       c("Untagged", "untagged", NA))
    nbrInt <- colSums(tstats$tstatsfilt[rownames(tstats$tstatsfilt) %in% all_tfs, ] > 0, 
                      na.rm = TRUE)
    names(nbrInt) <- .getSimplifiedComparison(names(nbrInt))

    ## Split by family - return a matrix with several columns: 
    ## #pulled down TFs (including itself), #pulled down TFs (excluding itself),
    ## #pulled down TFs from the same family (excluding itself), 
    ## #pulled down TFs from another family
    tmptf <- tstats$tstatsfilt[match(.getProteinNameFromComparison(colnames(tstats$tstatsfilt)), 
                                     rownames(tstats$tstatsfilt)), ]
    rownames(tmptf) <- .getProteinNameFromComparison(colnames(tstats$tstatsfilt))
    dbdtf <- data.frame(PomBaseID = .getPomBaseIdFromComparison(colnames(tstats$tstatsfilt), 
                                                                idmap = idmap),
                        bait = .getProteinNameFromComparison(colnames(tstats$tstatsfilt)),
                        DBD_class = NA)
    for (i in seq_len(nrow(dbdtf))) {
        lookfor <- dbdtf$PomBaseID[i]
        if (!(lookfor %in% c("untagged", "Untagged"))) {
            dbdtf$DBD_class[i] <- dbd$DBD_class[match(lookfor, dbd$PomBaseID)]
        } else {
            dbdtf$DBD_class[i] <- "no_class"
        }
    }
    stopifnot(dbdtf$bait == .getProteinNameFromComparison(colnames(tmptf)))
    nbrIntMat <- do.call(dplyr::bind_rows, lapply(seq_along(colnames(tmptf)), function(i) {
        expr <- colnames(tmptf)[i]
        fam <- dbdtf$DBD_class[i]
        data.frame(experiment = expr, 
                   bait = dbdtf$bait[i], 
                   dbdfam = fam, 
                   TFs = paste(rownames(tmptf)[-i][which(tmptf[-i, expr] > 0)],
                               collapse = ","),
                   nbrTFs = sum(tmptf[, expr] > 0, na.rm = TRUE),
                   nbrTFs_wobait = sum(tmptf[-i, expr] > 0, na.rm = TRUE),
                   nbrTFs_samefam = sum(tmptf[-i, expr] > 0 & 
                                            dbdtf$DBD_class[-i] == fam, 
                                        na.rm = TRUE),
                   nbrTFs_difffam = sum(tmptf[-i, expr] > 0 & 
                                            dbdtf$DBD_class[-i] != fam, 
                                        na.rm = TRUE)
        )
    }))

    ## Split experiments into three groups
    if (conc == 150) {
        stopifnot(all(colnames(tstat) == .getSimplifiedComparison(colnames(tstats$tstatsfilt))))
        tstat_no_bait <- 
            tstat[, colSums(tstats$tstatsfilt) == 0 | 
                      (.getOrigBaitNameFromComparison(colnames(tstats$tstatsfilt), idmap = idmap) %in% 
                      baitclass$Gene_name[baitclass$class == "nobait"])]
        tstat_150only <- 
            tstat[, colSums(tstats$tstatsfilt) > 0 &
                      (.getOrigBaitNameFromComparison(colnames(tstats$tstatsfilt), idmap = idmap) %in% 
                      baitclass$Gene_name[baitclass$class == "150only"])]
        tstat_150and500 <- 
            tstat[, colSums(tstats$tstatsfilt) > 0 &
                      (.getOrigBaitNameFromComparison(colnames(tstats$tstatsfilt), idmap = idmap) %in% 
                      baitclass$Gene_name[baitclass$class == "150and500"])]

        proteins_150only <- .getProteinNameFromSimplifiedComparison(colnames(tstat_150only))
        stopifnot(all(proteins_150only %in% rownames(tstat)))
        tstat_150and500 <- tstat_150and500[!(rownames(tstat_150and500) %in% proteins_150only), ]

        ## Cluster proteins based on the correlation between their t-stat vectors
        dst <- sqrt(2 * (1 - cor(t(tstat_150and500), use = "pairwise.complete")))
        dst[is.na(dst)] <- sqrt(2 * (1 - (-1)))
        clst <- hclust(as.dist(dst))

        ## Reorder proteins by the clustering
        protorder <- c(rownames(tstat_150and500)[clst$order], proteins_150only)
        tstat <- tstat[protorder, ]

        ## Reorder experiments to keep a "diagonal" in the heatmap
        tstat <- reorderColsByRows(tstat)

        tmpbaits <- .getOrigBaitNameFromSimplifiedComparison(colnames(tstat), 
                                                             idmap = idmap)
        colsplit <- ifelse(
            tmpbaits %in% baitclass$Gene_name[baitclass$class == "nobait"], 
            "No\nbait", 
            ifelse(
                tmpbaits %in% baitclass$Gene_name[baitclass$class == "150only"],
                "150 mM NaCl",
                ifelse(
                    tmpbaits %in% baitclass$Gene_name[baitclass$class == "150and500"],
                    "150 and\n500 mM NaCl", "NA"
                )
            )
        )
    } else if (conc == 500) {
        tstat_onlybait <- 
            tstat[, colSums(tstats$tstatsfilt > 0) == 1 |
                      (.getOrigBaitNameFromComparison(colnames(tstats$tstatsfilt), 
                                                      idmap = idmap) %in% 
                      baitclass$Gene_name[baitclass$class500 == "Lost in 500 mM"])]
        tstat_several <- 
            tstat[, colSums(tstats$tstatsfilt > 0) > 1 & 
                      (.getOrigBaitNameFromComparison(colnames(tstats$tstatsfilt), 
                                                      idmap = idmap) %in% 
                      baitclass$Gene_name[baitclass$class500 != "Lost in 500 mM"])]
        tstat_onlybait <- tstat_onlybait[, colnames(tstat_onlybait) != "Untagged_plate"]
        tstat_several <- tstat_several[, colnames(tstat_several) != "Untagged_plate"]

        proteins_onlybait <-
            .getProteinNameFromSimplifiedComparison(colnames(tstat_onlybait))
        stopifnot(all(proteins_onlybait %in% rownames(tstat)))
        tstat_several <- tstat_several[!(rownames(tstat_several) %in% proteins_onlybait), ]

        ## Cluster proteins based on the correlation between their t-stat vectors
        dst <- sqrt(2 * (1 - cor(t(tstat_several), use = "pairwise.complete")))
        dst[is.na(dst)] <- sqrt(2 * (1 - (-1)))
        clst <- hclust(as.dist(dst))

        ## Reorder proteins by the clustering
        protorder <- c(rownames(tstat_several)[clst$order], proteins_onlybait)
        tstat <- tstat[protorder, ]

        ## Reorder conditions to keep a "diagonal" in the heatmap
        tstat <- reorderColsByRows(tstat)

        colsplit <- factor(
            structure(baitclass$class500[
                match(.getOrigBaitNameFromSimplifiedComparison(colnames(tstat), 
                                                               idmap = idmap), 
                      baitclass$Gene_name)],
                names = colnames(tstat)),
            levels = c("Retained in 500 mM", "Lost in 500 mM", ""))
        levels(colsplit) <- c("Interactions\nin 500 mM NaCl",
                              "No interactions\nin 500 mM NaCl", "")
    } else {
        stop("Unknown concentration")
    }
    return(list(tstat = tstat, colsplit = colsplit, nbrIntMat = nbrIntMat))
}

makeBaitHeatmapData <- function(sce, adjpthr, log2fcthr, nbrIntMat) {
    tc <- .getTestCols(sce, adjp_cutoff = adjpthr, logfc_cutoff = log2fcthr)$tstat
    colnames(tc) <- .getSimplifiedComparison(colnames(tc))
    shared_names <- intersect(rownames(tc), 
                              .getProteinNameFromSimplifiedComparison(colnames(tc)))
    shared_names <- setdiff(shared_names, .getProteinFromOrigBait(nobait_150, idmap = idmap))
    stopifnot(all(shared_names %in% nbrIntMat$bait))
    nbrIntMat$complex <- NA
    for (cplx in c("MBF transcription complex", "CCAAT-binding factor complex", "Atf1-Pcr1")) {
        nbrIntMat$complex[.getPomBaseIdFromProtein(nbrIntMat$bait,
                                                   idmap = idmap) %in% 
                              complexes[[cplx]]$pombase_ids] <- cplx
    }
    shared_names <- nbrIntMat[match(shared_names, nbrIntMat$bait), ] |>
        arrange(dbdfam, complex) |>
        pull(bait)
    mat <- tc[shared_names, match(shared_names, .getProteinNameFromSimplifiedComparison(colnames(tc)))]
    rownames(nbrIntMat) <- .getSimplifiedComparison(nbrIntMat$experiment)
    list(mat = mat, nbrIntMat = nbrIntMat)
}
```

@@

## High stringency

Here we use an adjusted p-value cutoff of `r adjpThreshold` and a log2 
fold change cutoff of `r log2fcThreshold`.

@@

    data.frame(`Total number of TFs for which bait was pulled down` = ncol(ni_int),
               `Total number of interactions (excluding bait itself)` = length(which(ni_int > 0)),
               `Number of TFs with interactions` = length(which(colSums(ni_int, na.rm = TRUE) > 0)),
               `Number of TFs without observed interactions` = length(which(colSums(ni_int, na.rm = TRUE) == 0)),
               `Total number of TF-TF interactions (excluding bait itself)` = sum(baitdata$nbrIntMat$nbrTFs_wobait),
               `Number of TFs involved in these TF-TF interactions` = length(unique(c(tmp$bait, unlist(strsplit(tmp$TFs, ","))))),
               `Number of bilateral TF-TF interaction` = nbr_bilateral,
               check.names = FALSE) |>
        pivot_longer(cols = everything(), names_to = "Aspect", values_to = "Value") |>
        kbl(col.names = NULL, caption = tbl_title) |>
        kable_styling()
}

## 150mM
## --------------------------------------------------------------------------
makeInteractorSummary(sce = sce150, baitdata = baitdata_150, adjp_cutoff = adjpThreshold,
                      logfc_cutoff = log2fcThreshold, idmap = idmap, 
                      baits_to_keep = baitclass$Gene_name[baitclass$class != "nobait"],
                      tbl_title = "150 mM, high stringency")

## 500mM
## --------------------------------------------------------------------------
makeInteractorSummary(sce = sce500, baitdata = baitdata_500, adjp_cutoff = adjpThreshold,
                      logfc_cutoff = log2fcThreshold, idmap = idmap, 
                      baits_to_keep = baitclass$Gene_name[baitclass$class != "nobait"],
                      tbl_title = "500 mM, high stringency")
```

## Low stringency

Here we use an adjusted p-value cutoff of `r adjpThresholdLoose` and a log2 
fold change cutoff of `r log2fcThreshold`.

```{r}
## 150mM
## --------------------------------------------------------------------------
makeInteractorSummary(sce = sce150, baitdata = baitdata_150_loose, adjp_cutoff = adjpThresholdLoose,
                      logfc_cutoff = log2fcThreshold, idmap = idmap, 
                      baits_to_keep = baitclass$Gene_name[baitclass$class != "nobait"],
                      tbl_title = "150 mM, low stringency")

## 500mM
## --------------------------------------------------------------------------
makeInteractorSummary(sce = sce500, baitdata = baitdata_500_loose, adjp_cutoff = adjpThresholdLoose,
                      logfc_cutoff = log2fcThreshold, idmap = idmap, 
                      baits_to_keep = baitclass$Gene_name[baitclass$class != "nobait"],
                      tbl_title = "500 mM, low stringency")
```

# Figure 4

## Settings

```{r plot-settings}
colLabel_150 <- expand.grid(c("Untagged", "Ace2", "Rst2", "Pap1"),
                            c("tube", "plate")) |>
    tidyr::unite(col = "col", Var1, Var2) |>
    dplyr::pull(col)
colLabel_500 <- expand.grid(c("Untagged", "Atf1", "Moc3", "Ntu1",
                              "Pcr1", "Ntu2"),
                            c("tube", "plate")) |>
    tidyr::unite(col = "col", Var1, Var2) |>
    dplyr::pull(col)
rowLabel <- NULL

complexLabel <- c("NuA4", "SAGA")
```

@@

### Overall heatmap

```{r heatmap-tstat-150}
#| fig.width: 6
#| fig.height: 8

hm1a <- Heatmap(hmdata_150$tstat, 
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                col = makeHeatmapCol(hmdata_150$tstat, stringency = "high"), 
                border = TRUE, 
                border_gp = gpar(lwd = 0.5),
                column_split = hmdata_150$colsplit, 
                name = "Mod.\nt-stat.", 
                na_col = binary_heatmap_colors["FALSE"],
                column_names_gp = gpar(fontsize = fs),
                row_names_gp = gpar(fontsize = fs),
                row_title = "Proteins significantly enriched in at least one experiment",
                row_title_gp = gpar(fontsize = fl),
                show_row_names = FALSE, show_column_names = FALSE,
                use_raster = FALSE, show_heatmap_legend = TRUE,
                right_annotation = makeComplexAnnotation(
                    hmdata_150$tstat, complexes[complexLabel], 
                    idmap = idmap),
                bottom_annotation = makeColLabels(hmdata_150$tstat, colLabel_150, fl),
                width = unit(w_ipmsHm, "mm"), # heatmap body width
                heatmap_height = unit(h_ipmsHm, "mm"), # whole heatmap height
                column_title_gp = gpar(fontsize = ft),
                heatmap_legend_param = list(title_gp = gpar(fontsize = fl), 
                                border = "gray10",
                                border_gp = gpar(lwd = 0.5)))

set.seed(42L)
hm_150 <- grid.grabExpr(
    hm1a <- draw(hm1a, merge_legend = TRUE,
                heatmap_legend_side = "right",
                annotation_legend_side = "right"),
    width = w_ipmsHm, height = h_ipmsHm
)

plot_grid(hm_150)
```

@@

@@

@@

rnabinding <- read.delim(params$rnabinders)
keep <- rownames(hmdata_150$tstat) %in% c(rnabinding$PomBaseID, .capitalize(rnabinding$Gene_name))
sum(keep)
hm1rna <- Heatmap(hmdata_150$tstat[keep, ], 
                  cluster_rows = FALSE, 
                  cluster_columns = TRUE, 
                  col = makeHeatmapCol(hmdata_150$tstat, stringency = "high"), 
                  border = TRUE, 
                  border_gp = gpar(lwd = 0.5),
                  name = "Mod.\nt-stat.", 
                  na_col = binary_heatmap_colors["FALSE"],
                  column_names_gp = gpar(fontsize = fs),
                  row_names_gp = gpar(fontsize = fs),
                  show_row_names = TRUE, show_column_names = FALSE,
                  use_raster = FALSE, show_heatmap_legend = TRUE,
                  # bottom_annotation = makeColLabels(tstat150[keep, ], colLabel_150, fl),
                  #left_annotation = makeRowLabels(tstat150[keep, ], rowLabel, fl),
                  width = unit(w_ipmsHm, "mm"), # heatmap body width
                  heatmap_height = unit(h_ipmsHm/2, "mm"), # whole heatmap height
                  heatmap_legend_param = list(title_gp = gpar(fontsize = fl), 
                                border = "gray10",
                                border_gp = gpar(lwd = 0.5)))

@@

dnabinding <- read.delim(params$dnabinders)
keep <- rownames(hmdata_150$tstat) %in% c(dnabinding$PomBaseID, .capitalize(dnabinding$Gene_name))
sum(keep)
hm1dna <- Heatmap(hmdata_150$tstat[keep, ], 
                  cluster_rows = FALSE, 
                  cluster_columns = TRUE, 
                  col = makeHeatmapCol(hmdata_150$tstat, stringency = "high"), 
                  border = TRUE, 
                  border_gp = gpar(lwd = 0.5),
                  name = "Mod.\nt-stat.", 
                  na_col = binary_heatmap_colors["FALSE"],
                  column_names_gp = gpar(fontsize = fs),
                  row_names_gp = gpar(fontsize = fs),
                  show_row_names = TRUE, show_column_names = FALSE,
                  use_raster = FALSE, show_heatmap_legend = TRUE,
                  # bottom_annotation = makeColLabels(tstat150[keep, ], colLabel_150, fl),
                  #left_annotation = makeRowLabels(tstat150[keep, ], rowLabel, fl),
                  width = unit(w_ipmsHm, "mm"), # heatmap body width
                  heatmap_height = unit(1.5 * h_ipmsHm, "mm"), # whole heatmap height
                  heatmap_legend_param = list(title_gp = gpar(fontsize = fl), 
                                border = "gray10",
                                border_gp = gpar(lwd = 0.5)))

@@

### Overall heatmap

```{r heatmap-tstat-500}
#| fig.width: 6
#| fig.height: 8

hm2a <- Heatmap(hmdata_500$tstat, 
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                col = makeHeatmapCol(hmdata_500$tstat, stringency = "high"), 
                border = TRUE, 
                border_gp = gpar(lwd = 0.5),
                column_split = hmdata_500$colsplit, 
                name = "Mod.\nt-stat.", 
                na_col = binary_heatmap_colors["FALSE"],
                column_names_gp = gpar(fontsize = fs),
                row_names_gp = gpar(fontsize = fs),
                row_title = "Proteins significantly enriched in at least one experiment",
                row_title_gp = gpar(fontsize = fl),
                show_row_names = FALSE, show_column_names = FALSE,
                use_raster = FALSE, show_heatmap_legend = TRUE,
                right_annotation = makeComplexAnnotation(hmdata_500$tstat, complexes[complexLabel],
                                                         idmap = idmap, show_legend = FALSE),
                bottom_annotation = makeColLabels(hmdata_500$tstat, colLabel_500, fl),
                width = unit(w_ipmsHm, "mm"), # heatmap body width
                heatmap_height = unit(h_ipmsHm, "mm"), # whole heatmap height
                column_title_gp = gpar(fontsize = ft),
                heatmap_legend_param = list(title_gp = gpar(fontsize = fl), 
                                border = "gray10",
                                border_gp = gpar(lwd = 0.5)))

hm_500 <- grid.grabExpr(
    hm2a <- draw(hm2a, merge_legend = TRUE,
                 heatmap_legend_side = "right",
                 annotation_legend_side = "right"),
    width = w_ipmsHm, height = h_ipmsHm
)

plot_grid(hm_500)
```

@@

>>>>

# Helper function

```{r}
## Overwrite the get_legend function from cowplot temporarily, 
## since the cowplot one doesn't work with ggplot2 3.5.0
## see https://github.com/wilkelab/cowplot/issues/202

get_legend <- function(plot, legend = NULL) {
    gt <- ggplotGrob(plot)

    pattern <- "guide-box"
    if (!is.null(legend)) {
        pattern <- paste0(pattern, "-", legend)
    }

    indices <- grep(pattern, gt$layout$name)

    not_empty <- !vapply(
        gt$grobs[indices], 
        inherits, what = "zeroGrob", 
        FUN.VALUE = logical(1)
    )
    indices <- indices[not_empty]

    if (length(indices) > 0) {
        return(gt$grobs[[indices[1]]])
    }
    return(NULL)
}
```

@@

@@

## Summary tables

The 'high stringency' thresholds correspond to an adjusted p-value cutoff of 
`r adjpThreshold` and a log2 fold change cutoff of `r log2fcThreshold`.
The 'low stringency' thresholds correspond to an adjusted p-value cutoff of 
`r adjpThresholdLoose` and a log2 fold change cutoff of `r log2fcThreshold`.

@@

    data.frame(`Total number of TFs for which bait was pulled down` = ncol(ni_int),
               `Total number of interactions (excluding bait itself)` = length(which(ni_int > 0)),
               `Number of TFs with interactions` = length(which(colSums(ni_int, na.rm = TRUE) > 0)),
               `Number of TFs without observed interactions` = length(which(colSums(ni_int, na.rm = TRUE) == 0)),
               `Total number of TF-TF interactions (excluding bait itself)` = sum(baitdata$nbrIntMat$nbrTFs_wobait),
               `Number of TFs involved in these TF-TF interactions` = length(unique(c(tmp$bait, unlist(strsplit(tmp$TFs, ","))))),
               `Number of bilateral TF-TF interaction` = nbr_bilateral,
               check.names = FALSE) |>
        pivot_longer(cols = everything(), names_to = "Aspect", values_to = tbl_title)
}

makeInteractorSummary(sce = sce150, baitdata = baitdata_150, adjp_cutoff = adjpThreshold,
                      logfc_cutoff = log2fcThreshold, idmap = idmap, 
                      baits_to_keep = baitclass$Gene_name[baitclass$class != "nobait"],
                      tbl_title = "150 mM, high stringency") |>
    dplyr::full_join(
        makeInteractorSummary(sce = sce500, baitdata = baitdata_500, adjp_cutoff = adjpThreshold,
                              logfc_cutoff = log2fcThreshold, idmap = idmap, 
                              baits_to_keep = baitclass$Gene_name[baitclass$class != "nobait"],
                              tbl_title = "500 mM, high stringency")
    ) |> 
    dplyr::full_join(
        makeInteractorSummary(sce = sce150, baitdata = baitdata_150_loose, adjp_cutoff = adjpThresholdLoose,
                              logfc_cutoff = log2fcThreshold, idmap = idmap, 
                              baits_to_keep = baitclass$Gene_name[baitclass$class != "nobait"],
                              tbl_title = "150 mM, low stringency")
    ) |>
    dplyr::full_join(
        makeInteractorSummary(sce = sce500, baitdata = baitdata_500_loose, adjp_cutoff = adjpThresholdLoose,
                              logfc_cutoff = log2fcThreshold, idmap = idmap, 
                              baits_to_keep = baitclass$Gene_name[baitclass$class != "nobait"],
                              tbl_title = "500 mM, low stringency")
    ) |>
    kbl() |>
    kable_styling()
```

# Figure 5

@@

@@

```{r}
## Pie chart, within-vs-between DBD family interactions
pie_bait_150 <- ggplot(
    baitdata_150$nbrIntMat |>
        select(nbrTFs_samefam, nbrTFs_difffam) |>
        pivot_longer(everything(), names_to = "class", values_to = "nbrTF") |>
        group_by(class) |>
        summarize(nbrTF = sum(nbrTF), .groups = "drop"),
    aes(x = "", y = nbrTF, fill = class)) + 
    geom_col(width = 1) + 
    coord_polar("y", start = 0) + 
    scale_fill_manual(values = c(nbrTFs_samefam = main_colors[3],
                                 nbrTFs_difffam = main_colors[5]),
                      labels = c(nbrTFs_samefam = "Same DBD family",
                                 nbrTFs_difffam = "Different DBD family"),
                      name = "") + 
    labs(title = "150 mM NaCl\nhigh stringency") + 
    theme_void()
```

@@

```{r}
## Pie chart, within-vs-between DBD family interactions
pie_bait_150_loose <- ggplot(
    baitdata_150_loose$nbrIntMat |>
        select(nbrTFs_samefam, nbrTFs_difffam) |>
        pivot_longer(everything(), names_to = "class", values_to = "nbrTF") |>
        group_by(class) |>
        summarize(nbrTF = sum(nbrTF), .groups = "drop"),
    aes(x = "", y = nbrTF, fill = class)) + 
    geom_col(width = 1) + 
    coord_polar("y", start = 0) + 
    scale_fill_manual(values = c(nbrTFs_samefam = main_colors[3],
                                 nbrTFs_difffam = main_colors[5]),
                      labels = c(nbrTFs_samefam = "Same DBD family",
                                 nbrTFs_difffam = "Different DBD family"),
                      name = "") + 
    labs(title = "150 mM NaCl\nlow stringency") + 
    theme_void()
```

@@

rnabinding <- read.delim(params$rnabinders)
keep <- rownames(hmdata_150$tstat) %in% c(rnabinding$PomBaseID, .capitalize(rnabinding$Gene_name))
sum(keep)
hm1rna <- Heatmap(hmdata_150$tstat[keep, ], 
                  cluster_rows = FALSE, 
                  cluster_columns = TRUE, 
                  col = makeHeatmapCol(hmdata_150$tstat, stringency = "high"), 
                  border = TRUE, 
                  border_gp = gpar(lwd = 0.5),
                  name = "Mod.\nt-stat.", 
                  na_col = binary_heatmap_colors["FALSE"],
                  column_names_gp = gpar(fontsize = fs),
                  row_names_gp = gpar(fontsize = fs),
                  show_row_names = TRUE, show_column_names = FALSE,
                  use_raster = FALSE, show_heatmap_legend = TRUE,
                  width = unit(w_ipmsHm, "mm"), # heatmap body width
                  heatmap_height = unit(h_ipmsHm/2, "mm"), # whole heatmap height
                  heatmap_legend_param = list(title_gp = gpar(fontsize = fl), 
                                border = "gray10",
                                border_gp = gpar(lwd = 0.5)))

@@

dnabinding <- read.delim(params$dnabinders)
keep <- rownames(hmdata_150$tstat) %in% c(dnabinding$PomBaseID, .capitalize(dnabinding$Gene_name))
sum(keep)
hm1dna <- Heatmap(hmdata_150$tstat[keep, ], 
                  cluster_rows = FALSE, 
                  cluster_columns = TRUE, 
                  col = makeHeatmapCol(hmdata_150$tstat, stringency = "high"), 
                  border = TRUE, 
                  border_gp = gpar(lwd = 0.5),
                  name = "Mod.\nt-stat.", 
                  na_col = binary_heatmap_colors["FALSE"],
                  column_names_gp = gpar(fontsize = fs),
                  row_names_gp = gpar(fontsize = fs),
                  show_row_names = TRUE, show_column_names = FALSE,
                  use_raster = FALSE, show_heatmap_legend = TRUE,
                  width = unit(w_ipmsHm, "mm"), # heatmap body width
                  heatmap_height = unit(1.5 * h_ipmsHm, "mm"), # whole heatmap height
                  heatmap_legend_param = list(title_gp = gpar(fontsize = fl), 
                                border = "gray10",
                                border_gp = gpar(lwd = 0.5)))

@@

@@

```{r}
## Pie chart, within-vs-between DBD family interactions
pie_bait_500 <- ggplot(
    baitdata_500$nbrIntMat |>
        select(nbrTFs_samefam, nbrTFs_difffam) |>
        pivot_longer(everything(), names_to = "class", values_to = "nbrTF") |>
        group_by(class) |>
        summarize(nbrTF = sum(nbrTF), .groups = "drop"),
    aes(x = "", y = nbrTF, fill = class)) + 
    geom_col(width = 1) + 
    coord_polar("y", start = 0) + 
    scale_fill_manual(values = c(nbrTFs_samefam = main_colors[3],
                                 nbrTFs_difffam = main_colors[5]),
                      labels = c(nbrTFs_samefam = "Same DBD family",
                                 nbrTFs_difffam = "Different DBD family"),
                      name = "") + 
    labs(title = "500 mM NaCl\nhigh stringency") + 
    theme_void()
```

```{r}
## Pie chart, within-vs-between DBD family interactions
pie_bait_500_loose <- ggplot(
    baitdata_500_loose$nbrIntMat |>
        select(nbrTFs_samefam, nbrTFs_difffam) |>
        pivot_longer(everything(), names_to = "class", values_to = "nbrTF") |>
        group_by(class) |>
        summarize(nbrTF = sum(nbrTF), .groups = "drop"),
    aes(x = "", y = nbrTF, fill = class)) + 
    geom_col(width = 1) + 
    coord_polar("y", start = 0) + 
    scale_fill_manual(values = c(nbrTFs_samefam = main_colors[3],
                                 nbrTFs_difffam = main_colors[5]),
                      labels = c(nbrTFs_samefam = "Same DBD family",
                                 nbrTFs_difffam = "Different DBD family"),
                      name = "") + 
    labs(title = "500 mM NaCl\nlow stringency") + 
    theme_void()
```

## Pie charts for bait-vs-bait interactions (within/between DBD families)

```{r}
#| label: pie-bait-summary
#| fig.width: 10
#| fig.height: 3

cowplot::plot_grid(
    cowplot::plot_grid(pie_bait_150 + theme(legend.position = "none"), 
                       pie_bait_150_loose + theme(legend.position = "none"), 
                       pie_bait_500 + theme(legend.position = "none"),
                       pie_bait_500_loose + theme(legend.position = "none"), nrow = 1),
    get_legend(pie_bait_150 + theme(legend.position = "bottom")),
    ncol = 1, rel_heights = c(1, 0.25)
)
```
